---
title: "XShare"
author: "Luke Maurits"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r import data, include=FALSE}

### Import data
#### Load packages and clear environment.
library(tidyverse)
library(rstan)
library(matrixStats)
library(knitr)

d <- read_csv("../data/raw_data_anonymized.csv")

#Make color pallette
pal80s<-rep(c("#eab60b","#4d8d72","#ea84b2","#e67309","#516bb4","#e45d5c","#9c84e8","#7ed2e8","#e9b50d"),20)
```

## Raw data summaries and visualisation

```{r}
d %>%
  group_by(pid) %>% slice(1) %>% ungroup() %>%
  dplyr::summarise(n = n_distinct(pid),
		   n_children = sum(age<17),
		   n_adults = sum(age>=17),
		   min_age = min(age),
		   max_age=max(age),
		   mean_age=mean(age),
		   n_fem=sum(sex=="FEMALE"),
		   n_BaYaka=sum(culture=="BAYAKA"))

d %>%
  group_by(pid) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::mutate(
    age_group = ifelse(age < 17, "child", "adult"),
    age_bin = case_when(
      age < 2  ~ "0-1",
      age < 4  ~ "2-3",
      age < 6  ~ "4-5",
      age < 8  ~ "6-7",
      age < 10 ~ "8-9",
      age < 12 ~ "10-11",
      age < 14 ~ "12-13",
      age < 16 ~ "14-15",
      age < 20 ~ "17-19",
      age < 30 ~ "20-29",
      age < 40 ~ "30-39",
      age < 50 ~ "40-49",
      age < 60 ~ "50-59",
      age < 70 ~ "60-69",
      age < 80 ~ "70-79",
      TRUE     ~ "80+"
    ),
    age_bin = factor(age_bin, levels = c(
      "0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15",
      "17-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80+"
    ))
  ) %>%
  group_by(culture, age_bin, age_group,sex) %>%
  dplyr::summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = age_bin, y = n, fill = sex)) +
  geom_col() +
  facet_grid(~culture) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = element_blank()) +
  labs(x = "Age Bin", y = "Count")
```

## Model results (model with sex effects, but marginalising over them for reporting)

```{r}
model <- readRDS("model_sex.rds")
x <- extract(model)
```

### Sharing rates

The estimated ``infant'' sharing rate, with posterior mean `r round(mean(x$start_prob), 2)`, with 89% HPD interval [`r round(quantile(x$start_prob, p=0.055), 2)` - `r round(quantile(x$start_prob, p=0.945), 2)`].

``` {r}
# Marginalize over sex for adult sharing rates
nd <- expand_grid(culture = c(1, 2), condition = c(1, 2))
nd$index <- 2 * (nd$culture - 1) + nd$condition
nd$culture <- c("Bandongo", "Bayaka")[nd$culture]
nd$condition <- c("Inter-ethnic", "Intra-ethnic")[nd$condition]

for (i in 1:nrow(nd)) {
  # Compute male and female posteriors
  end_probs_male <- plogis(qlogis(x$end_probs[, nd$index[i]]) + x$end_prob_sex_effects[, nd$index[i]] * (-0.5))
  end_probs_female <- plogis(qlogis(x$end_probs[, nd$index[i]]) + x$end_prob_sex_effects[, nd$index[i]] * 0.5)
  
  # Average over sex (marginalization)
  end_probs_avg <- (end_probs_male + end_probs_female) / 2

  # Summarize posterior
  nd$posterior_mean[i] <- mean(end_probs_avg)
  nd$hpd_lower[i] <- quantile(end_probs_avg, p = 0.055)
  nd$hpd_upper[i] <- quantile(end_probs_avg, p = 0.945)
}

nd %>%
  dplyr::select(-index) %>%
  rename(`End probability posterior mean` = posterior_mean,
         `End probability 89% HPD lower` = hpd_lower,
         `End probability 89% HPD upper` = hpd_upper) %>%
  kable(digits = 2)
```

### Completion ages

```{r}
completion_age <- function(x, index) {
    result <- c()
    for(sex in c(-0.5, 0.5)) {
        result <- c(result, x$age_params[, index, 1] +
	                     x$age_param_1_sex_effects[, index]*sex -
                             log(0.05) * (x$age_params[, index, 2] +
					  x$age_param_2_sex_effects[, index] * sex))
    }
    return(result)
}

for (i in 1:nrow(nd)) {
  completion_ages <- completion_age(x, nd$index[i])
  nd$posterior_mean[i] <- mean(completion_ages)
  nd$hpd_lower[i] <- quantile(completion_ages, 0.055)
  nd$hpd_upper[i] <- quantile(completion_ages, 0.945)
}

nd %>%
  dplyr::select(-index) %>%
  rename(`Completion age posterior mean` = posterior_mean,
         `Completion age 89% HPD lower` = hpd_lower,
         `Completion age 89% HPD upper` = hpd_upper) %>%
  kable(digits = 2)
```

### Differences in completion ages

```{r}
contrasts <- tibble(culture=integer(), posterior_mean = numeric(), hpd_lower = numeric(),
	            hpd_upper = numeric(), support = numeric())
for(culture in c(1, 2)) {
	condition <- 1
	index <- 2*(culture - 1) + condition
	inter_completion_ages <- completion_age(x, index)
	condition <- 2
	index <- 2*(culture - 1) + condition
	intra_completion_ages <- completion_age(x, index)
	contrast <- inter_completion_ages - intra_completion_ages
	contrasts <- add_row(contrasts, culture=culture, posterior_mean = mean(contrast),
			     hpd_lower = quantile(contrast, p=0.055),
			     hpd_upper = quantile(contrast, p=0.945),
			     support = mean(contrast > 0))
}
contrasts$culture <- c("Bandongo", "Bayaka")[contrasts$culture]
contrasts %>%
  rename(`Completion age posterior mean` = posterior_mean,
         `Completion age 89% HPD lower` = hpd_lower,
         `Completion age 89% HPD upper` = hpd_upper,
	 `Posterior support for hypothesis` = support) %>%
  kable(digits = 2)
```

### Posterior predictive plots

```{r}
sharing_probability <- function(start_prob,
				end_prob, end_prob_sex_effect,
				age_param_1, age_param_1_sex_effect,
				age_param_2, age_param_2_sex_effect,
				age) {
    p <- 0
    for(sex in c(-0.5, 0.5)) {
        sex_end_prob <- plogis(qlogis(end_prob) + end_prob_sex_effect*sex)
        sex_age_param_1 <- age_param_1 + age_param_1_sex_effect*sex
        sex_age_param_2 <- age_param_2 + age_param_2_sex_effect*sex
        p <- p + 0.5*(start_prob +
	              (sex_end_prob - start_prob)*
		      plogis(-sex_age_param_2*(age - sex_age_param_1)))
    }
    return(p)
}

nd <- expand_grid(age=seq(5, 60, 0.25), culture=c(1, 2), condition=c(1, 2))
nd$index <- 2*(nd$culture - 1) + nd$condition

# First add posterior mean
start_prob <- mean(x$start_prob)
end_probs <- colMeans(x$end_probs)
end_prob_sex_effects <- colMeans(x$end_prob_sex_effects)
age_params <- colMeans(x$age_params)
age_param_1_sex_effects <- colMeans(x$age_param_1_sex_effects)
age_param_2_sex_effects <- colMeans(x$age_param_2_sex_effects)
for (i in 1:nrow(nd)) {
    nd$p[i] <- sharing_probability(start_prob,
                                   end_probs[nd$index[i]], end_prob_sex_effects[nd$index[i]],
                                   age_params[nd$index[i], 1], age_param_1_sex_effects[nd$index[i]],
                                   age_params[nd$index[i], 2], age_param_2_sex_effects[nd$index[i]],
                                   nd$age[i])
}
nd$type <- "posterior mean"
nd$alpha <- 1
nd$sample <- 0
plot_t <- nd

# Now plot 50 random posterior samples
nd$type <- "posterior samples"
nd$alpha <- 0.5
indices <- sample(1:length(x$start_prob), size=50, replace=FALSE)
for(j in indices) {
    start_prob <- x$start_prob[j]
    end_probs <- x$end_probs[j, ]
    end_prob_sex_effects <- x$end_prob_sex_effects[j, ]
    age_params <- x$age_params[j, , ]
    age_param_1_sex_effects <- x$age_param_1_sex_effects[j, ]
    age_param_2_sex_effects <- x$age_param_2_sex_effects[j, ]
    for (i in 1:nrow(nd)) {
        nd$p[i] <- sharing_probability(start_prob,
                                       end_probs[nd$index[i]], end_prob_sex_effects[nd$index[i]],
                                       age_params[nd$index[i], 1], age_param_1_sex_effects[nd$index[i]],
                                       age_params[nd$index[i], 2], age_param_2_sex_effects[nd$index[i]],
                                       nd$age[i])
    }
    nd$sample <- j
    plot_t <- add_row(plot_t, nd)
}

# Inflection ages
d_inf <- expand_grid(culture=c(1, 2), condition=c(1, 2))
d_inf$index <- 2*(d_inf$culture - 1) + d_inf$condition
d_inf$age <- 0
for(i in 1:nrow(d_inf)) {
	inf_ages <- 0.5*(x$age_params[,d_inf$index[i],1] + x$age_param_1_sex_effects[,d_inf$index[i]]*-0.5)
	inf_ages <- inf_ages + 0.5*(x$age_params[,d_inf$index[i],1] + x$age_param_1_sex_effects[,d_inf$index[i]]*0.5)
	d_inf$age[i] <- mean(inf_ages)
}
d_inf$culture <- c("Bandongo", "Bayaka")[d_inf$culture]
d_inf$condition <- c("Inter-ethnic", "Intra-ethnic")[d_inf$condition]

#  Do plot
plot_t$culture <- c("Bandongo", "Bayaka")[plot_t$culture]
plot_t$condition <- c("Inter-ethnic", "Intra-ethnic")[plot_t$condition]

plot_t <- plot_t %>%
  dplyr::mutate(condition = as.factor(condition),
         condition = fct_relevel(condition,"Intra-ethnic", "Inter-ethnic"))
d_inf <- d_inf %>%
  dplyr::mutate(condition = as.factor(condition),
         condition = fct_relevel(condition,"Intra-ethnic", "Inter-ethnic"))

ggplot(plot_t) +
	geom_line(aes(x=age, y=p, group=sample, colour=condition, linewidth=type, alpha=alpha)) +
	geom_vline(aes(xintercept=age), linetype="dotted", data=d_inf) +
	scale_discrete_manual("linewidth", values=c(1, 1/3)) +
	facet_grid(condition~culture) +
	scale_x_log10(breaks=c(5, 10, 20, 30, 40, 50),
		      minor_breaks=NULL,
		      limits=c(5, 60)) +
	scale_colour_discrete(name = "Condition",
			      type=c("#43a2ca","#feb24c")) +
	xlab("Age (years)") +
	ylab("Probability of sharing") +
	ylim(c(0.15,0.85)) +
	guides(alpha=FALSE, linewidth=FALSE)
ggsave("plot3.png", dpi=600, width=180, height=180, units="mm")
```
